title: "Business Intelligence Lab Submission Markdown"
author: "naive"
date: "4/10/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one student name, class group (just the letter; A, B, or C), and ID per line, e.g., 123456 - A - John Leposo; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  135575 - B - Dennis Nzioki.                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 2.  134645 - B - Vivean Lydiah                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 3.  134765 - B - Nicholas Munene                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 4.  131653- B - Terry Joan                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 5.  124428 - B - Eston Gichuhi                                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | *\<specify the name of the team you created on GitHub classroom\>*                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages
We start by installing all the required packages
```{r}

if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

# Introduction ----
# The performance of the trained models can be compared visually. This is done
# to help you to identify and choose the top performing models.

# STEP 1. Install and Load the Required Packages ----
## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## kernlab ----
if (require("kernlab")) {
  require("kernlab")
} else {
  install.packages("kernlab", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## randomForest ----
if (require("randomForest")) {
  require("randomForest")
} else {
  install.packages("randomForest", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

```
# STEP 2. Load the Dataset ----
```{r}
data("iris")

```
## 3.a. Train the Models ----
 We train the following models, all of which are using 10-fold repeated cross
 validation with 3 repeats:
 LDA
KNN
SVM
Random Forest
CART
```{r}



train_control <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

### LDA ----
set.seed(7)
iris_model_lda <- train(Species ~ ., data = iris,
                                method = "lda", trControl = train_control)


### KNN ----
set.seed(7)
iris_model_knn <- train(Species ~ ., data = iris,
                                method = "knn", trControl = train_control)

### SVM ----
set.seed(7)
iris_model_svm <- train(Species ~ ., data = iris,
                                method = "svmRadial", trControl = train_control)

### Random Forest ----
set.seed(7)
iris_model_rf <- train(Species ~ ., data = iris,
                               method = "rf", trControl = train_control)
### CART ----
set.seed(7)
iris_model_cart <- train(Species ~ ., data = iris,
                         method = "rpart", trControl = train_control)
## 3.b. Call the `resamples` Function ----
# We then create a list of the model results and pass the list as an argument
# to the `resamples` function.

results <- resamples(list(LDA = iris_model_lda, CART = iris_model_cart, 
                          KNN = iris_model_knn, SVM = iris_model_svm,
                          RF = iris_model_rf))
```
# STEP 4. Display the Results ----
this will give a visual representation of the key evaluation metrics for the models i.e. accuracy and Kappa
```{r}
## 1. Table Summary ----
# This is the simplest comparison. It creates a table with one model per row
# and its corresponding evaluation metrics displayed per column.

summary(results)

## 2. Box and Whisker Plot ----
# This is useful for visually observing the spread of the estimated accuracies
# for different algorithms and how they relate.

scales <- list(x = list(relation = "free"), y = list(relation = "free"))
bwplot(results, scales = scales)

## 3. Dot Plots ----
# They show both the mean estimated accuracy as well as the 95% confidence
# interval (e.g. the range in which 95% of observed scores fell).

scales <- list(x = list(relation = "free"), y = list(relation = "free"))
dotplot(results, scales = scales)

## 4. Scatter Plot Matrix ----
# This is useful when considering whether the predictions from two
# different algorithms are correlated. If weakly correlated, then they are good
# candidates for being combined in an ensemble prediction.

splom(results)

## 5. Pairwise xyPlots ----
# You can zoom in on one pairwise comparison of the accuracy of trial-folds for
# two models using an xyplot.

# xyplot plots to compare models
xyplot(results, models = c("LDA", "SVM"))

# or
# xyplot plots to compare models
xyplot(results, models = c("SVM", "RF"))

## 6. Statistical Significance Tests ----
# This is used to calculate the significance of the differences between the
# metric distributions of the various models.

### Upper Diagonal ----
# The upper diagonal of the table shows the estimated difference between the
# distributions. If we think that LDA is the most accurate model from looking
# at the previous graphs, we can get an estimate of how much better it is than
# specific other models in terms of absolute accuracy.

### Lower Diagonal ----
# The lower diagonal contains p-values of the null hypothesis.
# The null hypothesis is a claim that "the distributions are the same".
# A lower p-value is better (more significant).

diffs <- diff(results)

summary(diffs)

```




